---
title: "Code Challenges"
description: Additional materials for R Fundamentals
---

## Project Management

```{verbatim project_tree}

project_name
├── project_name.Rproj
├── README.md
├── script_1.R
├── script_2.R
├── data
│   ├── processed
│   └── raw
├── results
└── temp
```

## Data Structures

### Coercion Rules

```{verbatim coercion}

logical -> integer -> numeric -> complex -> character

```

### Challenge 1

1. Make a vector with numbers 1 - 26
2. Multiply the vector by 2
3. Name the vector elements A - Z (Hint: there is a built-in vector called `LETTERS`)
4. Print the last five items

### Challenge 2

You want to produce the following matrix:

```{verbatim matrix_wrapping}

     [,1] [,2]
[1,]    4    1
[2,]    9    5
[3,]   10    7
```

Which of the following is correct command to do that?

```{r matrix_creation_options}

matrix(c(4, 1, 9, 5, 10, 7), nrow = 3)
matrix(c(4, 9, 10, 1, 5, 7), ncol = 2, byrow = TRUE)
matrix(c(4, 9, 10, 1, 5, 7), nrow = 2)
matrix(c(4, 1, 9, 5, 10, 7), ncol = 2, byrow = TRUE)
```

## Exploring Data Frames

### Challenge 3

Create a new data frame from `gapminder` that includes the following columns:

1. country
2. year
3. gdpPercap

### Challenge 4

Create a new data frame that contains all of the 21st century data from European countries.

## Control Flow

### Conditional template
```{r conditional_statements}

# if
if (condition is true) {
    perform action
}

# if ... else
if (condition is true) {
    perform action
} else {  # if the condition is false,
    perform alternative action
}
```

### Iteration template

```{r for_loops}

for (iterator in set of values) {
  do a thing
}
```

## Functions

### Function template

```{r functions}

my_function <- function(parameters) {
  # perform action
  # return value
}
```

### Challenge 6

1. Calculate the total GDP for each North American country in 1952 and 1957
2. Save your results as a CSV file called `north_america_1.csv` in a subfolder called `data`.

